---
title: "Eggs Data Wrangling Solution"
author: "Sean Conway"
date: "8/1/2021"
output: pdf_document
---

```{r setup, include=FALSE,message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
```

## The Data
```{r echo=F, message=F}
eggs <- read_csv(here("R","data_cleaned","eggs_2004_2013.csv"))
```

```{r}
eggs
```
  
```{r}
glimpse(eggs)
```

## \textcolor{red}{DATA WRANGLING - SOLUTION}
    
### Creating `carton_size` & `price` variables 

1. Use `pivot_longer()` to combine the names of the egg carton sizes into a single variable, `carton_size`, while moving the values contained in these columns to another variable, `price`. 
  
#### Solution   
  
Using `pivot_longer()`, we can take our wide dataset and elongate it to make it tidy. First, we define the columns (`cols`) we want, using the colon operator (`:`) to select all four carton size columns. Next, we use `names_to` to take our column names and put them in a single column which we will call `carton_size`. Then, the price values, which were spread across columns, are now placed in a single column, called `price`.
 
```{r}
eggs_1 <- eggs %>%
  pivot_longer( 
    cols=large_half_dozen:extra_large_dozen, 
    names_to = "carton_size", 
    values_to = "price" 
  )
```

### Converting `price` to `price_dollar`
  
2. Use `mutate()` to convert `price` to dollar values, in a new variable called `price_dollar`. Drop `price` from the data.
  
#### Solution  
We use `mutate()` to convert `price` to `price_dollars`, and then remove the `price` column (which we no longer need) using `select(-c(price))`.
 
```{r}
eggs_1_dollar <- eggs_1 %>% 
  mutate(
    price_dollar=price/100
  ) %>%
  select(-c(price))
eggs_1_dollar
```

### Creating a `season` variable  
  
3. Use `mutate()` and `case_when()` to create a new column, `season`, based on the values of the column `month`. Here are the "rules" for this new column:
    - If `month` is equal to `"September"`, `"October"`, or `"November"`, `season` should have the value `"fall"`.
    - If `month` is equal to `"December"`, `"January"`, or `"February"`, `season` should have the value `"winter"`.
    - If `month` is equal to `"March"`, `"April"`, or `"May"`, `season` should have the value `"spring"`.
    - If `month` is equal to `"June"`, `"July"`, or `"August"`, `season` should have the value `"summer"`.
  
#### Solution
  
  
First, we need to create a new variable, called `season`, which classifies each month as being in winter, spring, summer, or fall. There are two solutions here (though other solutions are possible).
  
##### Solution 1 - The Long Way  
  
We use `mutate()` and `case_when()` to create a variable called `season`, where the value of `month` is used to classify each row as winter, spring, summer, or fall.
```{r}
eggs_tidy <- eggs_1_dollar %>%
  mutate( 
    season=case_when( 
      month == "December" | month == "January" | month == "February" ~ "winter",
      month =="March" | month == "April" | month == "May" ~ "spring",
      month == "June" | month == "July" | month == "August" ~ "summer",
      month == "September" | month == "October" | month == "November" ~ "fall"
    )
  )
```
  
###### Solution 1 - The results  
  
```{r}
eggs_tidy
```

##### Solution 2 - The Short(er) Way  
  
We can first create four vectors, `winter`, `spring`, `summer`, `fall`, each containing the names of the months in the particular season. 

Then, we use `mutate()`, `case_when()`, and the `%in%` operator to check if each row's `month` value is in each vector. For example, if the `month` in a particular row is `"January"`, the `case_when()` call will return a value of `TRUE`, and call it `winter`. This continues for each of the four seasons. 

```{r}
winter <- c("December","January","February")
spring <- c("March", "April", "May")
summer <- c("June","July","August")
fall <- c("September", "October","November")

eggs_tidy_1 <- eggs_1 %>%
  mutate(
    season=case_when(
      month %in% winter ~ "winter", 
      month %in% spring ~ "spring",
      month %in% summer ~ "summer",
      month %in% fall ~ "fall"
    )
  )
```

###### Solution 2 - The results  
  
  
```{r}
eggs_tidy_1
```
  
The results look the same, so we'll just use `eggs_tidy` for the remainder of the file.
 
## \textcolor{red}{DATA QUESTIONS - SOLUTION}
  
## Answer the following questions: 
  
1. How much did a large carton of a half-dozen eggs cost in October 2008?
2. Which month has the highest average price for a large carton of a half-dozen eggs (ignoring the year)?
3. Which year had the highest average price for a an extra large carton of a dozen eggs?
4. In 2009, which season (i.e., fall, winter, spring, summer) had the lowest average price for a large carton of a dozen eggs?
5. What was the median price for one extra-large carton of a half-dozen eggs in summer 2011?
  
### Question 1 - Solution
  
1. How much did a large carton of a half-dozen eggs cost in October 2008?
```{r}
eggs_tidy %>%
  filter(month=="October" & year==2004 & carton_size=="large_half_dozen") 
```
The answer is $1.28.  
  
### Question 2 - Solution
  
2. Which month has the highest average price for a large carton of a half-dozen eggs (ignoring the year)?

```{r}
eggs_tidy %>%
  filter(carton_size=="large_half_dozen") %>%
  group_by(month) %>%
  summarise(avg_price=mean(price_dollar)) %>%
  arrange(desc(avg_price))
```

It's a tie! between 7 months - June, July, August, December, September, October, November, with an average price of $1.57.
  
### Question 3 - Solution
  
3. Which year had the highest average price for a an extra large carton of a dozen eggs?

```{r}
eggs_tidy %>% 
  filter(carton_size=="extra_large_dozen") %>%
  group_by(year) %>%
  summarise(avg_price=mean(price_dollar)) %>%
  arrange(desc(avg_price))
```
2013 had the highest average price for extra large dozen eggs, at $2.90.

### Question 4 - Solution
  
4. In 2009, which season (i.e., fall, winter, spring, summer) had the lowest average price for a large carton of a dozen eggs?

```{r}
eggs_tidy %>%
  filter(carton_size=="large_dozen" & year==2009) %>%
  group_by(season) %>%
  summarise(avg_price=mean(price_dollar)) %>%
  arrange(desc(avg_price))
```
Spring is the answer, with an average price of $2.78.

### Question 5 - Solution
  
5. What was the median price for one extra-large carton of a half-dozen eggs in summer 2011?
  
```{r}
eggs_tidy %>%
  filter(carton_size=="extra_large_half_dozen" & year==2004 & season=="summer") %>%
  summarise(med_price=median(price_dollar))
```
The answer is $1.37. 
